MSK IV

### 2017-07-27 msk4.rmd
quick save

@@

>>>>

## Trigger 2

Kids more likely for acute osteomyelitis (haematogenous route) whereas adults more likely for chronic (direct).

Imaging to see osteomyelitis:

-  Xray: 2 weeks after infn
-  MRI: a couple of days

Questions to ask for a possible primary cause:

-  Any claudication
-  Check for diabetes
-  General inspection for malnutrition

## Trigger 3

Most common cause for osteomyelitis: *Staph*, *Pseudomonas* is stinky and common in diabetics,

-  Haematogenous: staph and influenza B (common in kids)
-  Open fractures: staph e coli
-  Spinal: staph, strep, e coli, IV drug users = candida
-  Diabetes/PVD: polymicrobial

Pathogenesis:

-  Infection
-  Inflammatory response
-  Phagocytosis attempt to contain infection by liberating enzymes
-  Enzymes lyse the bone
-  Pus forms
-  Pus travels through Haversian system
-  Increased intraosseous pressure
-  End arteritis and impaired blood flow
-  Ishaemic necrosis of bone
-  Pus breaks through cortex
-  Abscess
-  Subperiosteal infection
-  Elevated periosteum (involucrum) 
-  Rupture of periosteum(cloacae)

@@

Common causes of shoulder pain:

-  Glenohumeral
    -  RC muscles
    -  Bursae
    -  Tendons
    -  Adhesive capsulitis
    -  Labrum
    -  Synovium
    -  Articular cartilage
    -  Articular bone surface
-  Extraglenohumeral
    -  MSK: OA common in AC joint, diaphragmatic irritation
    -  CVS: MI
    -  C spine
    -  Malignancy: pancoast tumour (Horner's syndrome)
    -  Herpes
    -  PE
    -  Depression
    -  Polymyalgia

DDx: also RA, malignancy, tendinopathy (-itis),

Questions: Any particular movements that hurt? Have you started doing anything new recently? Do an ECG.

## Trigger 2

Hx indicative of pathology: probably acute, maybe traumatic

Overhead work: puts pressure on RC muscles

Abducting arms:

-  10-20: supraspinatus
-  20-90: deltoids
-  90+: levator scapulae

Smoking and shoulder pain:

-  Decreased bone density
-  Tendinopathy
-  Risk factor for RC tears

## Trigger 3
